Nail: A practical interface generator for data formats

### 2014-04-30 paper/osdi14/intro.tex
intro

Finally, data formats, for example PNG or PDF, contain features that are hard to represent with
existing parser generators. Some features, such as length and checksum fields, require the parser to
keep additional state. More complicated formats incorporate offset fields that require the parser to
reposition its input stream or consume the same input multiple times. In some cases, such as with
compressed data, very complicated transformations on the external format are necessary before it can
be understood. State-of-the-Art parser generators provide at most limited, unelegant facilities for
such features, so programmers are forced to manually writing control code to handle them, such as
moving the parsers input stream, looping over the invocation of a sub-parser or computing a checksum
over raw input bytes. Besides leaving much room for errors with offset arithmetic, such code is
usually not reusable when generating output.

This paper presents the initial design and implementation of Nail, a parser
generator that greatly reduces the programmer effort required to use a
grammar-based parser. Nail addresses the above four challenges with several key
ideas.

First, Nail reduces the expressiveness of its grammar language by removing
semantic actions. Existing parser generators allow arbitrary computation to
transform between the AST and the parser output. Instead, Nail derives the
structure of its output automatically from the grammar, forcing the programmer
to clearly separate syntactic validation and semantic processing.

>>>>

Finally, real-world data formats, for example PNG or PDF,  are hard to represent with
existing parser generators. Those parsers cannot directly deal with length or checksum fields
directly, so either the programmer has to write potentially unsafe code to deal with such features
or build contrived grammar constructs, such as introducing one grammar rule for each possible value
of a length fields. Offset fields usually require the programmer to manipulate a parser's internal
state to re-position its input. More complicated transformations on the external format, such as
compressed data, cannot be implemented at all.

This paper presents the design and implementation of Nail, a parser generator that greatly reduces
the programmer effort required to use grammars. Nail addresses the above four challenges with
several key ideas.

First, Nail grammars define both a format's external representation and an internal object model.
This removes the semantic actions and type declarations that programmers have to write with existing
parser generator. While this somewhat reduces the flexibility of the internal model, it forces the programmer
to clearly separate syntactic validation and semantic processing.
